Whose Truth is it Anyway? An Experiment on Annotation Bias

### 2023-06-08 report/draft.Rmd
update paper

# Whose Truth is it Anyway? Disagreement & Perspectivism in Creating Gold Standard Data
Generating large data-sets has become one of the main drivers of progress in natural language understanding.
In studies of political communication, the most familiar annotation tasks involve identifying basic concepts.
This often involve noting the topic of the text, the position of the actor or the tone of the text. 
A recent study [@van2021validity] demonstrates that using crowd-coding platforms is a good way to collect such large data-sets.
However, having only a few workers annotate the majority of text of interest has raised concerns about data diversity and models' ability to generalize beyond the crowd-workers:
In a series of experiments, @geva2019we show that often models do not generalize well to annotations from annotators that did not contribute to the training set, suggesting that annotator bias should be monitored during data-set creation.
One such potential bias, especially in times of increasing polarization [@iyengar2019origins; @gidron2019toward; @boxell2022cross], is based on ideological position of the annotator. 
Given that annotators on crows-coding platforms tend to be younger and hold more liberal values than the general public [@clifford2015samples;@huff2015these], this could potentially hamper the data diversity and generalizability of the model.
An additional reason to monitor the annotators ideological position as a potential source of annotator bias is that a recent study in NLP showed that experiential factors influence the consistency of how political ideologies are perceived [@shen2021sounds].
Their finding challenges the "ground-truth" assumption we as researcher make that a position for example is either left-leaning or right-wing leaning. 
People with different ideological backgrounds might experience that position differently. 
This challenges us.
We are interested in the effect of e.g. elite communication. 
To study this, we allow for heterogeneous treatment effects in experimental work.
This indicates that we often do not assume that the treatment, often using text, has the same effect for different partisans.
Yet, at the same time, we forget or ignore that knowledge when creating large data-sets for our machine learning models. 
To test whether people with different ideological backgrounds as well as their politicall knowledge might experience that position differently, challenging our ground-truth assumption in data annotation, we propose the following hypotheses:

@@

@@

If these biases exists, how can we alleviate them?
For the Austrian National Elections, @ennser2018impact already demonstrated that showing party labels impact annotators' assessment of the party position.
So, one solution is masking:

>>>>

# Whose Truth is it Anyway? Disagreement & Perspectivism in Creating Gold Standard Data
Generating large data-sets has become one of the main drivers of progress in natural language understanding.
In studies of political communication, the most familiar annotation tasks involve identifying basic concepts.
This often involve noting the topic of the text, the position of the actor or the tone of the text. 
A recent study [@van2021validity] demonstrates that using crowd-coding platforms is a good way to collect such large data-sets.
However, having only a few workers annotate the majority of text of interest has raised concerns about data diversity and models' ability to generalize beyond the crowd-workers:
In a series of experiments, @geva2019we show that often models do not generalize well to annotations from annotators that did not contribute to the training set, suggesting that annotator bias should be monitored during data-set creation.
One such potential bias, especially in times of increasing polarization [@iyengar2019origins; @gidron2019toward; @boxell2022cross], is based on ideological position of the annotator. 
Given that annotators on crows-coding platforms tend to be younger and hold more liberal values than the general public [@clifford2015samples;@huff2015these], this could potentially hamper the data diversity and generalizability of the model.
An additional reason to monitor the annotators ideological position as a potential source of annotator bias is that a recent study in NLP showed that experiential factors influence the consistency of how political ideologies are perceived [@shen2021sounds].
Their finding challenges the "ground-truth" assumption we as researcher make that a position for example is either left-leaning or right-wing leaning. 
People with different ideological backgrounds might experience that position differently. 
This challenges us.
We are interested in the effect of e.g. elite communication. 
To study this, we allow for heterogeneous treatment effects in experimental work.
This indicates that we often do not assume that the treatment, often using text, has the same effect for different partisans.
Yet, at the same time, we forget or ignore that knowledge when creating large data-sets for our machine learning models.

[Myrthe over NLP]

To test whether people with different ideological backgrounds as well as their political knowledge might experience that position differently, challenging our ground-truth assumption in data annotation, we propose the following hypotheses:

@@

[Myrthe over NLP]

To test whether people with different ideological backgrounds as well as their political knowledge might experience that position differently, challenging our ground-truth assumption in data annotation, we propose the following hypotheses:

@@

If these biases exists, how can we alleviate them?

[Myrthe over NLP]

For the Austrian National Elections, @ennser2018impact already demonstrated that showing party labels impact annotators' assessment of the party position.
So, one solution is masking:
